Examples Compendium

### 2021-01-13 examples.tex
multiplicity example

Now take $Z$ to be the MV cycle of weight $\alpha_{1,3}$ with Lusztig datum $(010,00,1)$ and leave $Z'$ as above. Then 
\begin{align}
    \young(13,2,4) \ast \young(11,23) &= \young(1113,223,4) + \young(1113,224,3) \\
    (1 \leftarrow 2 \to 3) \ast S_2 &= (2\to 3) \oplus (1\leftarrow 2) + P_2 \\
    (010,00,1) \ast (000,10,0) &= (010,10,1) + (010,01,0)
\end{align}
TODO: fill in the ideals.

@@

For this example, I think we need $\lambda_1 = (2,0,0)$, $\lambda_2 = (2,2,0)$, $\mu_1 = (0,2,0)$, $\mu_2 = (2,0,2)$. 
\[
\yng(2) = \{\young(22)\} \qquad \yng(2,2) = \{\young(11,33)\} 
\]
Then 
\[
A = \left[\begin{BMAT}(e){cc;cc;cc}{cc;cc;cc} 
    0 & 1 & 0 & 0 & 0 & 0\\
    -s^2 & 2s & A_0 & A_1 & A_2 & A_3\\
    0 & 0 & 0 & 1 & 0 & 0\\
    0 & 0 & 0 & 0 & A_4 & A_5\\
    0 & 0 & 0 & 0 & 0 & 1\\
    0 & 0 & 0 & 0 & -s^2 & 2s
    \end{BMAT}\right]
\]
needs $A-s\big|_{\CC^2\cap E_s}$ has Jordan type $(2)$, $A\big|_{\CC^4\cap E_0}$ has Jordan type $(2)$, and $A-s\big|_{E_s}$ has Jordan type $(2,2)$. 
So 
\[
E_0 = \{A_0 \e_1 + s^2\e_3\leftarrow A_0 \e_2 + s^2 \e_4 + (A_0 + A_1 + \frac {2A_0}  s  ) \e_1 + s^2 \e_3 \}   
\]
while 
\begin{gather}
E_s^1 = \{\e_1 + s \e_2\leftarrow \e_1 + (1 + s) \e_2 \}  \\
E_s^2 = \{ -\frac{A_2 + sA_3}{A_0 + sA_1}\e_3 - \frac{sA_2 + s^2 A_3}{A_0 + sA_1} \e_4 + \e_5 + s \e_6  \leftarrow\}
\end{gather}
if 
\begin{gather}
    0 = A_4 + sA_5 \\
    \to A_0 A_4 \text{ as } s \to 0 
\end{gather}
and

% Adding 2020-12-30 14:03:05: 
Let's also try  
\[
\begin{aligned}
    \lambda_2 &= (4,0,0) & \lambda_1 &= (4,4,0) & \lambda &= (8,4,0)\\
    \mu_2 &= (2,2,0)     & \mu_1 &= (4,2,2)     & \mu &= (6,4,2)
\end{aligned}    
\]
Then 
\[
\left[\begin{BMAT}(e){cccccc;cccc;cc}{cccccc;cccc;cc} 
    0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
    0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
    0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
    0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
    0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0\\
    0 & 0 & 0 & 0 & -s^2 & 2s & A_0 & A_1 & A_2 & A_3 & A_4 & A_5\\
    0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0\\
    0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0\\
    0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0\\
    0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & -s^2 & 2s & A_6 & A_7\\
    0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1\\
    0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0
    \end{BMAT}\right]    
\]
Note that there is only one SSYT of shape $\lambda_i$ and weight $\mu_i$
\[
\tau_2 = \young(1122) \qquad \tau_1 = \young(1111,2233)    
\]
Note also that 
$$\begin{aligned}
    t^4(t-s)^2 &= t^4 (t^2 - 2st + s^2) = t^6 - 2st^5 + s^2 t^4 \\
    t^2 (t-s)^2 &= t^4 - 2st^3 + s^2 t^2  % \\
    % t^2 &= t^2 
\end{aligned}$$
Elements of $\TT_{\mu_1,\mu_2}^+$ will take the form 
\[
A = \left[\begin{BMAT}(e){cccccc:cccc:cc}{cccccc:cccc:cc}
    & 1 & & & & & & & & & & \\
    & & 1 & & & & & & & & & \\
    & & & 1 & & & & & & & & \\
    & & & & 1 & & & & & & & \\
    & & & & & 1 & & & & & & \\
    & & & & -s^2 & 2s & 0 & 0 & a_3 & a_4 & b_1 & b_2 \\
    & & & & & & & 1 & & & & \\
    & & & & & & & & 1 & & & \\
    & & & & & & & & & 1 & & \\
    & & & & & & & & -s^2 & 2s & c_1 & c_2 \\
    & & & & & & & & & & & 1 \\
    & & & & & & & & & & & 
\end{BMAT}\right]
\]
The tableau tells us for each $1\le i\le 12$
\[
    \text{Jordan type of } A\big|_{\Sp(e_1,\dots,e_i)\cap E_0} \text{ is shape of }\tau_1\big|_{\text{first $i$ boxes}}
\]
So take $i = 10$. Then $A$ restricted to $\CC^{10} \cap E_0$ should have Jordan type $(4,2)$. 
\acom{How to do it box by box? $s$ columns somehow correspond to $\tau_2$ boxes.}
Therefore $a_1 = 0$. The 4-cycle is obvious. For the 2-cycle we require $a_2 = 0$. 
\[
    \begin{aligned}
e_1 &\leftarrow e_2 \leftarrow e_3 \leftarrow e_4 \\
e_7 &\leftarrow e_8 \\
% e_{11} &\leftarrow e_{12} 
    \end{aligned}    
\]
Now looking at all of $A$ we have continue our 2-cycle to a 4-cycle. Roger found 
\[
    e_{10} - \left(\frac{2s}{c_1}  + \frac{s^2 c_2}{c_1^2}\right) e_{11} + \frac{s^2}{c_1} e_{12}
\]
This requires $c_1^2 a_4 + c_1b_2s^2 - 2sc_1b_1-s^2 c_2 = 0$ and $a_3 c_1 + s^2 b_1 = 0$.

@@

\[
A-s = \left[\begin{BMAT}(e){cccccc:cccc:cc}{cccccc:cccc:cc}
    -s & 1 & & & & & & & & & & \\
    & -s & 1 & & & & & & & & & \\
    & & -s & 1 & & & & & & & & \\
    & & & -s & 1 & & & & & & & \\
    & & & & -s & 1 & & & & & & \\
    & & & & -s^2 & s & 0 & 0 & a_3 & a_4 & b_1 & b_2 \\
    & & & & & & -s & 1 & & & & \\
    & & & & & & & -s & 1 & & & \\
    & & & & & & & & -s & 1 & & \\
    & & & & & & & & -s^2 & s & c_1 & c_2 \\
    & & & & & & & & & & -s & 1 \\
    & & & & & & & & & & & -s 
\end{BMAT}\right]
\]
\end{example}

>>>>

Now take $Z$ to be the MV cycle of weight $\alpha_{1,3}$ with Lusztig datum $(010,00,1)$ and leave $Z'$ as above. Then 
\begin{align}
    \young(13,2,4) \ast \young(11,23) &= \young(1113,223,4) + \young(1113,224,3) \\
    (1 \leftarrow 2 \to 3) \ast S_2 &= (2\to 3) \oplus (1\leftarrow 2) + P_2 \\
    (010,00,1) \ast (000,10,0) &= (010,10,1) + (010,01,0)
\end{align}
% TODO: fill in the ideals. 
Verifying. Consider 
\[
    A = \left[\begin{BMAT}(e){ccc;cc;cc;c}{ccc;cc;cc;c} 
        0 & 1 & 0 & 0 & 0 & 0 & 0 & 0\\
        0 & 0 & 1 & 0 & 0 & 0 & 0 & 0\\
        0 & -s^2 & 2s & A_0 & A_1 & A_2 & A_3 & A_4\\
        0 & 0 & 0 & 0 & 1 & 0 & 0 & 0\\
        0 & 0 & 0 & 0 & s & A_5 & A_6 & A_7\\
        0 & 0 & 0 & 0 & 0 & 0 & 1 & 0\\
        0 & 0 & 0 & 0 & 0 & 0 & s & A_8\\
        0 & 0 & 0 & 0 & 0 & 0 & 0 & 0
        \end{BMAT}\right]    
\]
and 
\[
A - s = \left[\begin{BMAT}(e){ccc;cc;cc;c}{ccc;cc;cc;c} 
    -s & 1 & 0 & 0 & 0 & 0 & 0 & 0\\
    0 & -s & 1 & 0 & 0 & 0 & 0 & 0\\
    0 & -s^2 & s & A_0 & A_1 & A_2 & A_3 & A_4\\
    0 & 0 & 0 & -s & 1 & 0 & 0 & 0\\
    0 & 0 & 0 & 0 & 0 & A_5 & A_6 & A_7\\
    0 & 0 & 0 & 0 & 0 & -s & 1 & 0\\
    0 & 0 & 0 & 0 & 0 & 0 & 0 & A_8\\
    0 & 0 & 0 & 0 & 0 & 0 & 0 & -s
    \end{BMAT}\right]
\]
Then 
\begin{align}
    E_0 &= \{\e_1, \\
        & \e_4 \text{ if } A_0 = 0, \\
        &  s\e_6 - A_5 \e_5 + c\e_2  \text{ if } A_2 = 0 , \\
        & A_5 \e_8 - A_7 \e_6 \text{ if } A_4, A_8 = 0  \} \\ 
        &\to \{ \e_1 , \e_4, -A_5 \e_5 + c\e_2 , A_5 \e_8 - A_7 \e_6 \\
        &\text{if } A_0, A_2, A_4, A_8 = 0 \} \\
    E_s &= \{s^2 \e_3 + s \e_2 + \e_1, s\e_3 - \frac 1 s \e_1, \\
        & s\e_5 + \e_4 \text{ if } A_1 = 0, \\
        & s\e_7 + \e_6 + \e_5 \text{ if } A_5 + (A_6 - 1)s = 0 \} \label{ln:a3} \\
        &\to \{\e_1, \e_4, \e_6 + \e_5 \text{ if } A_1, A_5 = 0 \}
\end{align}
In line~\ref{ln:a3} I don't think $A_3 =0$ because we can probably add some combination $a\e_1 + b\e_2 + c\e_3$ so that $\e_7$ maps to both 1-cycles (the 2-d kernel) determined so far.

Thus $Z \ast Z'$ is contained in $(A_0,A_2,A_4,A_8, A_1, A_5)$

@@

% For this example, I think we need
Take $\lambda_1 = (2,0,0)$, $\lambda_2 = (2,2,0)$, $\mu_1 = (0,2,0)$, $\mu_2 = (2,0,2)$. Note that there is only one tableau of weight $\mu_i$ and type $\lambda_i$ for each $i$. 
\[
\yng(2) = \{\young(22)\} \qquad \yng(2,2) = \{\young(11,33)\} 
\]
Consider
\[
A = \left[\begin{BMAT}(e){cc;cc;cc}{cc;cc;cc} 
    0 & 1 & 0 & 0 & 0 & 0\\
    -s^2 & 2s & A_0 & A_1 & A_2 & A_3\\
    0 & 0 & 0 & 1 & 0 & 0\\
    0 & 0 & 0 & 0 & A_4 & A_5\\
    0 & 0 & 0 & 0 & 0 & 1\\
    0 & 0 & 0 & 0 & -s^2 & 2s
    \end{BMAT}\right]
\]
Then 
\begin{align}
E_s &= \{ \e_1 + s\e_2, \e_1 + (s+1) \e_2 \\
    & \e_5 + s\e_6 + \frac 1 s (A_4 + A_5 s) \e_4 + \frac 1 {s^2} (A_4 + A_5 s) \e_3 \\
    & \text{if } (A_0 + A_1 s) (A_4 + A_5s) + (A_2 + A_3 s) s^2 = 0 , \\
    & \e_1 + (s+a) \e_2 + \e_6 + \frac 1 s \e_5 + \frac 1 s A_5 \e_4 - \frac 1 {s^3} A_4 \e_3 \}  \\
    & \text{if } -2\,{A}_{3}s^{3}-{A}_{2}s^{2}+{A}_{0}{A}_{4}-{A}_{5}s = 0  \\
E_0 &= \{A_0 \e_1 + s^2 \e_3 , A_0 \e_2 + s^2 \e_4 + \frac{A_1}{A_0}\e_1 - 2s \e_3 \}
\end{align}
Taking $s\to 0$ we get 
\begin{align}
    E_{s=0} &= \{
        \e_1 , \e_1 + \e_2, \e_3 \text{ if } A_0A_4 = 0\} \\
    E_0 &= \{\e_1 , A_0^2 \e_2 + A_1 \e_1 \}
\end{align}
So $Z\ast Z'$ is contained in $(A_0A_4)$. If we look at $(A\big|_{s=0})^4$, which must be zero since $\lambda = (4,2)$, we pick up the additional equation ${A}_{1}{A}_{4}+{A}_{0}{A}_{5}$. 
Therefore the ideal of $Z\ast Z'$ is 
\begin{equation}
    (A_0A_4 , A_1A_4 + A_0 A_5 ) \subset \CC[A_0..A_5] 
\end{equation}
which decomposes as 
\((A_4, A_5) \sqcup (A_0, A_1) \sqcup (A_0,A_4) \). Since $$\CC[x,y]/(xy,x+y)\cong \CC[z]/(z^2)$$ the component $(A_0,A_4)$ occurs with multiplicity 2 as expected! Localize at, or colon out $(A_1,A_5)$.

Note that these ideals correspond to expected tableaux as follows.
\begin{align}
\young(22) \ast \young(11,33) &= \young(1122,33) + {\color{red}2?}\young(1123,23) + \young(1133,22)  \\
  &= (A_4, A_5) + {\color{red}2?}(A_0,A_4) + (A_0, A_1) \\
  &= (20,2) + {\color{red}2?}(11,1) + (02,0) \\
  &= (11\to 22) + {\color{red}2?}((1\to (1\leftarrow 2))\to 2) + (11\leftarrow 22) \\ 
  &= (xy-z)^2 + 2(xy-z)z + z^2
\end{align} 
\begin{question}
    How do we tell from the tableaux that an ideal is occuring with multiplicity? 
\end{question}

% Adding 2020-12-30 14:03:05: 
% Removing 2021-01-12 18:59:05
\begin{comment}
Let's also try  
\[
\begin{aligned}
    \lambda_2 &= (4,0,0) & \lambda_1 &= (4,4,0) & \lambda &= (8,4,0)\\
    \mu_2 &= (2,2,0)     & \mu_1 &= (4,2,2)     & \mu &= (6,4,2)
\end{aligned}    
\]
Then 
\[
\left[\begin{BMAT}(e){cccccc;cccc;cc}{cccccc;cccc;cc} 
    0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
    0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
    0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
    0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
    0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0\\
    0 & 0 & 0 & 0 & -s^2 & 2s & A_0 & A_1 & A_2 & A_3 & A_4 & A_5\\
    0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0\\
    0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0\\
    0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0\\
    0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & -s^2 & 2s & A_6 & A_7\\
    0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1\\
    0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0
    \end{BMAT}\right]    
\]
Note that there is only one SSYT of shape $\lambda_i$ and weight $\mu_i$
\[
\tau_2 = \young(1122) \qquad \tau_1 = \young(1111,2233)    
\]
Note also that 
$$\begin{aligned}
    t^4(t-s)^2 &= t^4 (t^2 - 2st + s^2) = t^6 - 2st^5 + s^2 t^4 \\
    t^2 (t-s)^2 &= t^4 - 2st^3 + s^2 t^2  % \\
    % t^2 &= t^2 
\end{aligned}$$
Elements of $\TT_{\mu_1,\mu_2}^+$ will take the form 
\[
A = \left[\begin{BMAT}(e){cccccc:cccc:cc}{cccccc:cccc:cc}
    & 1 & & & & & & & & & & \\
    & & 1 & & & & & & & & & \\
    & & & 1 & & & & & & & & \\
    & & & & 1 & & & & & & & \\
    & & & & & 1 & & & & & & \\
    & & & & -s^2 & 2s & 0 & 0 & a_3 & a_4 & b_1 & b_2 \\
    & & & & & & & 1 & & & & \\
    & & & & & & & & 1 & & & \\
    & & & & & & & & & 1 & & \\
    & & & & & & & & -s^2 & 2s & c_1 & c_2 \\
    & & & & & & & & & & & 1 \\
    & & & & & & & & & & & 
\end{BMAT}\right]
\]
The tableau tells us for each $1\le i\le 12$
\[
    \text{Jordan type of } A\big|_{\Sp(e_1,\dots,e_i)\cap E_0} \text{ is shape of }\tau_1\big|_{\text{first $i$ boxes}}
\]
So take $i = 10$. Then $A$ restricted to $\CC^{10} \cap E_0$ should have Jordan type $(4,2)$. 
\acom{How to do it box by box? $s$ columns somehow correspond to $\tau_2$ boxes.}
Therefore $a_1 = 0$. The 4-cycle is obvious. For the 2-cycle we require $a_2 = 0$. 
\[
    \begin{aligned}
e_1 &\leftarrow e_2 \leftarrow e_3 \leftarrow e_4 \\
e_7 &\leftarrow e_8 \\
% e_{11} &\leftarrow e_{12} 
    \end{aligned}    
\]
Now looking at all of $A$ we have continue our 2-cycle to a 4-cycle. Roger found 
\[
    e_{10} - \left(\frac{2s}{c_1}  + \frac{s^2 c_2}{c_1^2}\right) e_{11} + \frac{s^2}{c_1} e_{12}
\]
This requires $c_1^2 a_4 + c_1b_2s^2 - 2sc_1b_1-s^2 c_2 = 0$ and $a_3 c_1 + s^2 b_1 = 0$.

@@

\[
A-s = \left[\begin{BMAT}(e){cccccc:cccc:cc}{cccccc:cccc:cc}
    -s & 1 & & & & & & & & & & \\
    & -s & 1 & & & & & & & & & \\
    & & -s & 1 & & & & & & & & \\
    & & & -s & 1 & & & & & & & \\
    & & & & -s & 1 & & & & & & \\
    & & & & -s^2 & s & 0 & 0 & a_3 & a_4 & b_1 & b_2 \\
    & & & & & & -s & 1 & & & & \\
    & & & & & & & -s & 1 & & & \\
    & & & & & & & & -s & 1 & & \\
    & & & & & & & & -s^2 & s & c_1 & c_2 \\
    & & & & & & & & & & -s & 1 \\
    & & & & & & & & & & & -s 
\end{BMAT}\right]
\]
\end{comment}
\end{example}
